Limited information strategies for topological games

### 2015-03-09 phd/talks-and-presentations/2015-03-11-defense.tex
almost complete first draft of defense slides

  In doing so I uncovered several new results in general topology,
  advancing research done by
  G. Gruenhage, P. Nyikos, F. Gavin, R. Telg{\' a}rksy, J. Bell,
  M. Scheepers, and others.
\end{frame}

@@

@@

% \subsection{Proximal Game}

% \begin{frame}
%   \term{Proximal Game} (2014)
%   \cite{MR3239205}

%   \vspacing

%   \begin{figure}
%     \includegraphics[width=0.8\linewidth]{proximalGame.pdf}
%   \end{figure}
%   {\tiny for compact $T_1$ $0$-dim spaces}

%   \vpause

%   The first player $\pl D$ wins the game if the points chosen by the
%   second player $\pl P$ converge. If $\pl D$ has a winning
%   strategy for this game, call $X$ \term{proximal}.
% \end{frame}

% \begin{frame}
%   Some results related to the Proximal Game due to Jocelyn Bell:

%   \begin{proposition}
%     If $X$ is metrizable, then $X$ is proximal.
%   \end{proposition}

%   \begin{theorem}
%     If $X$ is proximal, then $X$ is collectionwise normal.
%   \end{theorem}

%   \begin{theorem}
%     $\Sigma$-products and closed subspaces of proximal spaces are proximal.
%   \end{theorem}

%   \pause

%   \begin{corollary}
%     The $\Sigma$-product of metrizable spaces is collectionwise normal.
%     \cite{MR0461410} \cite{MR716576}
%   \end{corollary}
% \end{frame}

% \section{Main Result}

% \subsection{Corson compact}

% \begin{frame}
%   A \term{Corson compact} space is a space homeomorphic to a compact
%   subset of the $\Sigma$-product of real lines.

%   \vpause

%   Peter Nyikos observed:

%   \begin{proposition}
%     Every Corson compact space is proximal compact. \cite{nyikosProximalPreprint}
%   \end{proposition}

%   \pause

%   C. and Gruenhage showed in \cite{MR3227201} that any compact proximal
%   space must be Corson compact, using another game-theoretic characterization
%   of Corson compact due to Gruenhage:
% \end{frame}

% \subsection{Showing proximal compact implies Corson compact}

% \begin{frame}
%   \small
%   \term{Diagonal Game} (1984) \cite{MR752278}:

%   \begin{figure}
%     \includegraphics[width=0.7\linewidth]{diagonalGame.pdf}
%   \end{figure}
%   {\tiny for compact $T_1$ $0$-dim spaces}

%   \vpause

%   The first player $\pl O$ wins the game if any open set containing the
%   diagonal also contains infinitely many of $\pl P$'s chosen points.

%   \begin{theorem}
%     For $X$ compact, $\pl O$ has a winning strategy for the diagonal game
%     if and only if $X$ is Corson compact.
%   \end{theorem}
% \end{frame}

% \begin{frame}
%   One may use a winning strategy $\sigma$ for $\pl D$ in the proximal game to
%   construct a strategy $\tau$ for $\pl O$ in the diagonal game.

%   \begin{figure}
%     \includegraphics[width=0.7\linewidth]{convertingStrategy.pdf}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   In general:
%   \[
%     \tau(a)
%       =
%     \bigcup_{s\concat\<i,h_{s,i},j\>\in\max(T(a))}
%     \frac{1}{4}\sigma(o_s\concat\<h_{s,i}\>)[h_{s,i,j}]
%   \]

%   \vspacing

%   Using the strategy $\tau$ defined for every proximal compact space,
%   $\pl O$ cannot be defeated in the diagonal game, and therefore all
%   proximal compacts are Corson compact. \qed
% \end{frame}

% \section{Questions and References}

% \begin{frame}
%   Open questions:

%   \begin{itemize}
%     \item If compactness is dropped, does the proximal game characterize
%     all copies of \textit{closed} subspaces of a $\Sigma$-product of reals?
%     (Nyikos)
%     \item If the winning strategy for the proximal game is \term{Markov}
%     (relies on only the latest move and round number) for a compact space,
%     does that imply that the space is \term{Eberlein} compact?
%     (This holds for the diagonal game.)
%   \end{itemize}
% \end{frame}

>>>>

  In doing so I uncovered several new results in general topology,
  advancing research done by
  G. Gruenhage, P. Nyikos, R. Telg{\' a}rksy, J. Bell,
  M. Scheepers, and others.
\end{frame}

@@

\section{Bell's Proximal Game}

\begin{frame}
  \small
  \begin{game}
  Bell's proximal game $\bellAbsConGame{X}$ for compact zero-dimensional
  $X$:
    \begin{figure}
      \includegraphics[width=0.6\linewidth]{proximalGameAlt.pdf}
    \end{figure}

  $\pl D$ wins the game if the points chosen by $\pl P$ converge.
  Otherwise, $\pl P$ wins.
  \end{game}
\end{frame}

\begin{frame}
  If $\pl D\win \bellAbsConGame{X}$, then $X$ is called a proximal compact.
  This game was brought to my attention due to this result: \cite{MR3239205}

  \begin{theorem}
    Every proximal space is a $W$-space. So
    $\pl D\win\bellAbsConGame{X} \Rightarrow \pl O\win\gruConGame{X}{x}$
    for all $x\in X$.
  \end{theorem}

  Proximal spaces have strong preservation properties, as any
  closed subset or $\Sigma$-product of proximal spaces is proximal. Since
  any proximal space is collectionwise normal, Bell's game gives an elegant
  proof of the classic result:

  \begin{theorem}
    A $\Sigma$-product of metrizable spaces is collection-wise normal.
  \end{theorem}
\end{frame}

\begin{frame}
  Nyikos \cite{nyikosProximalPreprint} observed that
  \begin{proposition}
    Corson compact spaces are proximal.
  \end{proposition}
  and asked if the converse holds as well.

  With Gruenhage, I showed that the answer is yes:
  \begin{theorem}
    A compact space is Corson compact if and only if it is proximal.
  \end{theorem}
\end{frame}

\begin{frame}\small
  A winning limited information strategy may always be passed down
  to win in a closed subspace, but Bell's result that winning strategies
  are preserved for $\Sigma$-products does not quite generalize as well:

  \begin{theorem}
    For $k<\omega$, if $\pl D\kmarkwin{k}\bellAbsConGame{X_i}$ for all $i<\omega$,
    then $\pl D\kmarkwin{k}\bellAbsConGame{\prod_{i<\omega} X_i}$.
  \end{theorem}

  Other limited information lemmas proved in my dissertation allowed me
  to prove a game-theoretic characterization of another compactness property
  (paper in preparation):

  \begin{theorem}
    A compact space is strong Eberlein compact if and only if
    $\pl D\tactwin\bellAbsConGame{X}$.
  \end{theorem}
\end{frame}

@@

\begin{frame}
  Gruenhage used these games in \cite{MR752278} to characterize
  metacompactness and $\sigma$-metacompactness amongst locally compact
  spaces:

  \begin{theorem}
    For locally compact spaces, $\pl K\tactwin\gruKPGame{X}$ if and only if
    $X$ is metacompact.
  \end{theorem}

  \begin{theorem}
    For locally compact spaces, $\pl K\markwin\gruKPGame{X}$ if and only if
    $X$ is $\sigma$-metacompact.
  \end{theorem}
\end{frame}

\begin{frame}
  By removing knowledge of the round number, an analogous result is
  unsurfaced:

  \begin{theorem}
    For locally compact spaces, $\pl K\prewin\gruKPGame{X}$ if and only if
    $X$ is $\sigma$-compact.
  \end{theorem}

  Actually, for locally compact or even compactly-generated spaces,
  $\pl K\prewin\gruKPGame{X}$ if and only if $\pl K\prewin\gruKLGame{X}$.
  However, there is a non-compactly-generated counterexample:

  \begin{theorem}
    There exists a free ultrafilter $\mc F$ such that
    $\pl K \prewin\gruKPGame{\omega\cup\{\mc F\}}$, but
    $\pl K \notprewin\gruKLGame{\omega\cup\{\mc F\}}$ for any free
    ultrafilter $\mc F$.
  \end{theorem}
\end{frame}

\section{Menger's Game}

\begin{frame}
  \small
  \begin{game}
  Menger's game $\menGame{X}$ proceeds as follows:

  (this space intentionally left blank until I make a picture)

  $\pl F$ wins the game if her finitely coverable subsets
  union to the space. Otherwise, $\pl C$  wins.
  \end{game}
\end{frame}

\begin{frame}
  A covering property generalizing $\sigma$-compactness is characterized
  by this game, demonstrated by Hurewicz in the 1920's. \cite{MR1544773}

  \begin{theorem}
    A space is Menger if and only if $\pl C\not\win\menGame{X}$.
  \end{theorem}

  It was originally suspected that Menger subspaces of the real line were
  exactly the $\sigma$-compact subspaces, but as was shown by Telgarsky
  and Scheepers, $\sigma$-compact spaces
  have slightly more structure. \cite{MR753073} \cite{MR1273523}

  \begin{theorem}
    A metrizable space $X$ is $\sigma$-compact if and only if
    $\pl F\win\menGame{X}$.
  \end{theorem}
\end{frame}

\begin{frame}
  By considering Markov strategies, the previous theorem may be factored
  into two subresults.

  \begin{theorem}
    A regular space $X$ is $\sigma$-compact if and only if
    $\pl F\markwin\menGame{X}$.
  \end{theorem}
  \begin{theorem}
    For a second-countable space $X$, $\pl F\win\menGame{X}$ if and only if
    $\pl F\markwin\menGame{X}$.
  \end{theorem}

  Note that since the spaces we are considering are all Lindel\"of,
  metrizability is characterized by regularity and second-countability.
\end{frame}

\subsection{More Motivation}

\begin{frame}
  Limited information strategies in topological games such as
  $\menGame{X}$ often have set-theoretic
  consequences. The statement $\alcompS\kappa$ due to M. Scheepers
  \cite{MR1129143} says that there exist
  almost-compatible functions $f_A:A\to\omega$ for each $A\in[\kappa]^\omega$.

  \begin{theorem}
    $\alcompS{\omega_1}$ and $\neg\alcompS{(2^\omega)^+}$ are theorems
    of $ZFC$, but $\alcompS\kappa$ is independent of $ZFC$ for
    $\omega_1<\kappa\leq 2^\omega$.
  \end{theorem}

  \begin{theorem}
    If $\alcompS{\kappa}$ holds, then
    $\pl F\ktactwin2 \cloFillStrictGame{\kappa}$.
  \end{theorem}
\end{frame}

\subsection{More Results}

\begin{frame}
  Let $\oneptlind\kappa$ be the one-point ``Lindel\"of-ication'' of
  discrete $\kappa$.

  \begin{theorem}
    \begin{figure}
      \includegraphics[width=0.6\linewidth]{mengerGameChart.png}
    \end{figure}
  \end{theorem}
\end{frame}

\begin{frame}\small
  For most of the games in the previous chart, including the Menger game
  $\menGame{X}$, there's no need to consider larger amounts of limited
  information.

  \begin{theorem}
    For each $k<\omega$,
    $\pl F\kmarkwin{k+2}\menGame{X}$ if and only if
    $\pl F\kmarkwin2\menGame{X}$
  \end{theorem}

  The topological property $\pl F\kmarkwin2\menGame{X}$ seems to
  depend on the set-theoretic axioms at play.

  \begin{theorem}
    If $\alcompS{2^\omega}$, then $\pl F\kmarkwin2\menGame{R_\omega}$.
  \end{theorem}
\end{frame}
